---
title: "p8105_hw6_sp4170"
author: "Shihui Peng"
date: "2023-12-02"
output: github_document
---
load packages and basic settings..
```{r, message=FALSE, echo=FALSE}
library(tidyverse)
library(modelr)
set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("data/homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```




# Problem 2

## load Central Park weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

## question 1: log(beta1_hat∗beta2_hat)

i first fit a linear regression model with y = tmax and covariants including tmin and prcp for my created 5000 samples (using bootstrap in modelr package). Then i use broom::tidy and unnest() to get estimates for tmins and prcps, select relevant columns and unnest the list. I use pivot_wider to allow the calculation between different rows within one column (that is, calculate log of products of estimates of tmin and estiamtes of prcps). then, i use mutate to calculate log(beta1_hat x beta2_hat), and draw a density plot based on the estimates of log(beta1_hat x beta2_hat). then, i calculate the 95%CI for log(beta1_hat x beta2_hat) estimates.

### fit linear regression and get log(beta1_hat∗beta2_hat) estimates

```{r}
boot_result_para =
  bootstrap(weather_df, 5000, id = "strap_number") |> 
  mutate(
    model = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    result_para = map(model, broom::tidy)
  ) |> 
  select(strap_number, result_para) |> 
  unnest(result_para) |> 
  select(strap_number, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  mutate(
    log_b1b2 = log(tmin * prcp)
  )
```

### draw a density plot for log(beta1_hat*beta2_hat) distribution

```{r}
boot_result_para |> 
  ggplot(aes(x = log_b1b2)) + geom_density() +
  labs(x = "log(beta1_hat*beta2_hat)")
```
descriptions: 

* the distribution of the log(beta1_hat x beta2_hat) is highly left-skewed.
* the peak of the log(beta1_hat x beta2_hat) distribution is around -5.8, and the majority of the estimated log(beta1_hat x beta2_hat) value lies between -8.2 and -4.4.(observed by eye)
* it is worth noticing that some values of beta1_hat x beta2_hat is negative, so we cannot calculate the log of them. in this case, this part of data is removed when we draw the density plot for estimated log(beta1_hat x beta2_hat).
  * for more details, `r boot_result_para |> filter(log_b1b2 != "NaN") |> count()` out of `r boot_result_para |> count()` rows are used when drawing the plot.

### calculate 95%CI for r^2 estimates

```{r}
para_ci =
boot_result_para |> 
  mutate(
    log_b1b2_lower_ci = quantile(log_b1b2, 0.025, na.rm = TRUE),
    log_b1b2_upper_ci = quantile(log_b1b2, 0.975, na.rm = TRUE)
  ) |> 
  select(log_b1b2_lower_ci, log_b1b2_upper_ci) |> 
  distinct() 
```
* based on the output, we can tell the 95% confidence interval for log(beta1_hat x beta2_hat), which i use 2.5% and 97.5% quantiles to get, is from `r para_ci |> pull(log_b1b2_lower_ci)` to `r para_ci |> pull(log_b1b2_upper_ci)`


## question 2: r^2

i first fit a linear regression model with y = tmax and covariants including tmin and prcp for my created 5000 samples using bootstrap() in modelr package. then, i use broom::glance to give r^2 values, select relevant columns and unnest the list. then, i draw a density plot based on the estimates of r^2. then, i calculate the 95%CI for r^2 estimates.

### fit linear regression and get r^2 estimates

```{r}
boot_result_r2 =
  bootstrap(weather_df, 5000, id = "strap_number") |> 
  mutate(
    model = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    result_r2 = map(model, broom::glance)
  ) |> 
  select(strap_number, result_r2) |> 
  unnest(result_r2) 
```

### draw a density plot for r^2 distribution

```{r}
boot_result_r2 |> 
  ggplot(aes(x = r.squared)) + geom_density() 
```
descriptions: 

* the distribution of r^2 is slightly left-skewed.
* the peak of the r^2 distribution is around 0.92, and the majority of the estimated r^2 value lies between 0.885 and 0.943. (observed by eye)

### calculate 95%CI for r^2 estimates

```{r}
r2_ci =
boot_result_r2 |> 
  mutate(
    r2_lower_ci = quantile(r.squared, 0.025),
    r2_upper_ci = quantile(r.squared, 0.975)
  ) |> 
  select(r2_lower_ci, r2_upper_ci) |> 
  distinct()
```
* based on the output, we can tell the 95% confidence interval for r^2, which i use 2.5% and 97.5% quantiles to get, is from `r r2_ci |> pull(r2_lower_ci)` to `r r2_ci |> pull(r2_upper_ci)`
* since the 95%CI is not wide and with relatively large value, we might say the model can make good prediction when we do bootstrap to infer r^2


# Problem 3

## data import, clean, and mutate

```{r}
bw_df = 
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(
    babysex = factor(babysex, levels = c(1,2), labels = c("male", "female")),
    malform = case_when(malform == 0 ~ "absent", malform == 1 ~ "present"),
    mrace = factor(mrace, levels = c(1,2,3,4,8), labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    frace = factor(frace, levels = c(1,2,3,4,8,9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknow"))
  )

colSums(is.na(bw_df))
```

* i import the data, clean the names, and mutate my interested columns for further steps. I drop NULL in bwt column as this is our response variable. then, i check the missing value in other columns, and find that there are no missing data.

## fit my model

* step 1
  * from my opinion, i suppose the main effect of (1) presence of malformations that could affect weight (malform), (2) mother’s race (mrace), (3) father’s race (frace), (4) average number of cigarettes smoked per day during pregnancy (smoken), (5) gestational age in weeks (gaweeks), and (6) baby’s head circumference at birth (bhead), will affect the baby’s birth weight. therefore, i fit a mixed linear regression model `fit1`
  * but i find that the r^2 is pretty low, so i decide to add a new covariant.
* step 2
  * i decide to add one more covariant (7) baby’s length at birth (blength) into the model `fit1`, and the new model is `fit2`
  * r^2 greatly increase, so i decide to keep variable `blength` into my mode.
* step 3
  * based on step 2 parameter estimates output, i find that the variable `malform` is not significant. so, i remove it from the model, and find that the r^2 does not change. To keep my model less complex, i decide to remove the variable `malform`, and here it comes my final model `fit`: `bwt ~ mrace + frace + smoken + gaweeks + blength`
```{r}
fit1 = lm(bwt ~ malform + mrace + frace + smoken + gaweeks + bhead, data = bw_df)

broom::tidy(fit1) |> knitr::kable()
broom::glance(fit1) |> knitr::kable()

fit2 = lm(bwt ~ malform + mrace + frace + smoken + gaweeks + bhead + blength, data = bw_df)

broom::tidy(fit2) |> knitr::kable()
broom::glance(fit2) |> knitr::kable()

my_fit = lm(bwt ~ mrace + frace + smoken + gaweeks + blength + bhead, data = bw_df)

broom::tidy(my_fit) |> knitr::kable()
broom::glance(my_fit) |> knitr::kable()
```

## show a plot of model residuals against fitted values based on my model `fit`

```{r}
bw_df |> 
  select(bwt, mrace, frace, smoken, gaweeks, blength, bhead) |> 
  add_predictions(my_fit) |> 
  add_residuals(my_fit) |> 
  ggplot(aes(y = resid, x = pred)) + geom_point(alpha = 0.5) +
  labs(x = "fitted value", y = "residuals", title = "residuals against fitted values, based on model `my_fit`")

```

* based on this output plot, we can see that there are lots of points around x-axis (0 line of the residuals axis), which indicates a relatively good prediction when using my model. however, when fitted value increases, we can see a trend that the absolute value of residuals kind of increases, so there might be a heteroscedasticity problem for residuals, which indicates the variance of residuals is not constant across the dataset.
* however, we can see some fitted values have large residuals, so maybe there are outliers when we use model fit for prediction. we need to take a further step to see whether these outliers are influential observations for this model.


## compare my model `fit` with other two models

### build & have a look at the other 2 models

i build two models. model `main_fit` is using length at birth and gestational age as predictors (main effects only). model `inter_fit` is using head circumference, length, sex, and all interactions (including the three-way interaction) between these. just take a look of these two models:
```{r}
main_fit = lm(bwt ~ blength + gaweeks, data = bw_df)

interaction_fit = lm(bwt ~ bhead * babysex * blength, data = bw_df)
```

### comparison between these three models

i use crossv_mc to create training sets and testing sets, and then convert resamples into tibbles.
```{r}
cv_df =
  bw_df |> 
  crossv_mc(n = 100) |> 
  mutate(
    train = map(train, as.tibble),
    test = map(test, as.tibble)
  )
```

i apply each model to the training set dataframe, and evaluate all testing set dataframe using root mean squared error (rmse).
```{r}
cv_result =
  cv_df |>
  mutate(
    my_fit = map(train, \(df) lm(bwt ~ mrace + frace + smoken + gaweeks + blength + bhead, data = df)),
    main_fit = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    interaction_fit = map(train, \(df) lm(bwt ~ bhead * babysex * blength, data = df))
  ) |> 
  mutate(
    rmse_my_fit = map2_dbl(my_fit, test, \(model, df) rmse(model, df)),
    rmse_main_fit = map2_dbl(main_fit, test, \(model, df) rmse(model, df)),
    rmse_interaction_fit = map2_dbl(interaction_fit, test, \(model, df) rmse(model, df))
  )
```

i draw a violin plot to show the root mean squared error for these three models..
```{r}
cv_result |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = 'model_type',
    values_to = 'rmse',
    names_prefix = 'rmse_'
  ) |> 
  ggplot(aes(x = model_type, y = rmse)) + geom_violin() +
  labs(
    title = "comparison among my proposed model `my_fit` and 2 given models",
    y = "root mean squared error"
  )
```

* based on the violin plot comparing my model and the other 2 given models, we can see:
  * the rmse of my model is less than the main_fit model without overlaps. theredfore, my proposed model is better than the main_fit model using length at birth and gestational age as predictors (main effects only) due to the smaller rmse. so the improvement of my model is pretty clear.
  * the rmse of my model is less than the interaction_fit mode, but the part of the distributions have overlapped. therefore, my proposed model might be better than the interaction_fit model using head circumference, length, sex, and all interactions, but the improvement is not really clear.
  