---
title: "p8105_hw6_sp4170"
author: "Shihui Peng"
date: "2023-12-02"
output: github_document
---
load packages and basic settings..
```{r, message=FALSE, echo=FALSE}
library(tidyverse)
library(modelr)
library(ggplot2)
set.seed(1)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


# Problem 2

## load Central Park weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

## draw 5000 samples from weather_df

I first create a function for sampling rows with size as large as 100% of the weather_df from weather_df. Then i iterate and create 5000 samples from weather_df using this function.
```{r}
boot_samp = function(df){
  sample_frac(df, replace = TRUE)
}

boot_strap =
  tibble(
    strap_number = 1:5000
  ) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_samp(weather_df))
  )
```

## question 1: log(beta1_hat∗beta2_hat)

i first fit a linear regression model with y = tmax and covariants including tmin and prcp for my created 5000 samples. Then i use broom::tidy and unnest() to get estimates for tmins and prcps, select relevant columns and unnest the list. I use pivot_wider to allow the calculation between different rows within one column (that is, calculate log of products of estimates of tmin and estiamtes of prcps). then, i use mutate to calculate log(beta1_hat x beta2_hat), and draw a density plot based on the estimates of log(beta1_hat x beta2_hat). then, i calculate the 95%CI for log(beta1_hat x beta2_hat) estimates.

### fit linear regression and get log(beta1_hat∗beta2_hat) estimates

```{r}
boot_result_para =
  boot_strap |> 
  mutate(
    model = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    result_para = map(model, broom::tidy)
  ) |> 
  select(strap_number, result_para) |> 
  unnest(result_para) |> 
  select(strap_number, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  mutate(
    log_b1b2 = log(tmin * prcp)
  )

# or this method????

boot_result_para =
  bootstrap(weather_df, 5000, id = "strap_number") |> 
  mutate(
    model = map(strap, \(df) lm(tmax ~ tmin + prcp, data = df)),
    result_para = map(model, broom::tidy)
  ) |> 
  select(strap_number, result_para) |> 
  unnest(result_para) |> 
  select(strap_number, term, estimate) |> 
  pivot_wider(
    names_from = term,
    values_from = estimate
  ) |> 
  mutate(
    log_b1b2 = log(tmin * prcp)
  )
```

### draw a density plot for log(beta1_hat*beta2_hat) distribution

```{r}
boot_result_para |> 
  ggplot(aes(x = log_b1b2)) + geom_density() +
  labs(x = "log(beta1_hat*beta2_hat)")
```
descriptions: 

* the distribution of the log(beta1_hat x beta2_hat) is highly left-skewed.
* the peak of the log(beta1_hat x beta2_hat) distribution is around -5.5, and the majority of the estimated log(beta1_hat x beta2_hat) value lies between -8.2 and -4.2.
* it is worth to notice that some values of beta1_hat x beta2_hat is negative, so we cannot calculate the log of them. in this case, this part of data is removed when we draw the density plot for estimated log(beta1_hat x beta2_hat).
  * for more details, `r boot_result_para |> filter(log_b1b2 != "NaN") |> count()` out of `r boot_result_para |> count()` rows are used when drawing the plot.

### calculate 95%CI for r^2 estimates

```{r}
para_ci =
boot_result_para |> 
  mutate(
    log_b1b2_lower_ci = quantile(log_b1b2, 0.025, na.rm = TRUE),
    log_b1b2_upper_ci = quantile(log_b1b2, 0.975, na.rm = TRUE)
  ) |> 
  select(log_b1b2_lower_ci, log_b1b2_upper_ci) |> 
  distinct() 
```
* based on the output, we can tell the 95% confidence interval for log(beta1_hat x beta2_hat), which i use 2.5% and 97.5% quantiles to get, is from `r para_ci |> pull(log_b1b2_lower_ci)` to `r para_ci |> pull(log_b1b2_upper_ci)`


## question 2: r^2

i first fit a linear regression model with y = tmax and covariants including tmin and prcp for my created 5000 samples. then, i use broom::glance to give r^2 values, select relevant columns and unnest the list. then, i draw a density plot based on the estimates of r^2. then, i calculate the 95%CI for r^2 estimates.

### fit linear regression and get r^2 estimates

```{r}
boot_result_r2 =
  boot_strap |> 
  mutate(
    model = map(strap_sample, \(df) lm(tmax ~ tmin + prcp, data = df)),
    result_r2 = map(model, broom::glance)
  ) |> 
  select(strap_number, result_r2) |> 
  unnest(result_r2) 
```

### draw a density plot for r^2 distribution

```{r}
boot_result_r2 |> 
  ggplot(aes(x = r.squared)) + geom_density() 
```
descriptions: 

* the distribution of r^2 is slightly left-skewed.
* the peak of the r^2 distribution is around 0.916, and the majority of the estimated r^2 value lies between 0.886 and 0.932.

### calculate 95%CI for r^2 estimates

```{r}
r2_ci =
boot_result_r2 |> 
  mutate(
    r2_lower_ci = quantile(r.squared, 0.025),
    r2_upper_ci = quantile(r.squared, 0.975)
  ) |> 
  select(r2_lower_ci, r2_upper_ci) |> 
  distinct()
```
* based on the output, we can tell the 95% confidence interval for r^2, which i use 2.5% and 97.5% quantiles to get, is from `r r2_ci |> pull(r2_lower_ci)` to `r r2_ci |> pull(r2_upper_ci)`
* since the 95%CI is not wide and with relatively large value, we might say the model can make good prediction when we do bootstrap to infer r^2



# Problem 3

## data import, clean, and mutate

```{r}
bw_df = 
  read_csv("data/birthweight.csv") |> 
  janitor::clean_names() |> 
  mutate(
    babysex = factor(babysex, levels = c(1,2), labels = c("male", "female")),
    malform = case_when(malform == 0 ~ "absent", malform == 1 ~ "present"),
    mrace = factor(mrace, levels = c(1,2,3,4,8), labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    frace = factor(frace, levels = c(1,2,3,4,8,9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknow"))
  ) |> 
  drop_na(bwt)

colSums(is.na(bw_df))
```

* i import the data, clean the names, and mutate my interested columns for further steps. I drop NULL in bwt column as this is our response variable. then, i check the missing value in other columns, and find that there are no missing data.

## fit my model

* step 1
  * from my opinion, i suppose the main effect of (1) presence of malformations that could affect weight (malform), (2) mother’s race (mrace), (3) father’s race (frace), (4) average number of cigarettes smoked per day during pregnancy (smoken), (5) gestational age in weeks (gaweeks), and (6) baby’s head circumference at birth (bhead), will affect the baby’s birth weight. therefore, i fit a mixed linear regression model `fit1`
  * but i find that the r^2 is pretty low, so i decide to add a new covariant.
* step 2
  * i decide to add one more covariant (7) baby’s length at birth (blength) into the model `fit1`, and the new model is `fit2`
  * r^2 greatly increase, so i decide to keep variable `blength` into my mode.
* step 3
  * based on step 2 parameter estimates output, i find that the variable `malform` is not significant. so, i remove it from the model, and find that the r^2 does not change. To keep my model less complex, i decide to remove the variable `malform`, and here it comes my final model `fit`: `bwt ~ mrace + frace + smoken + gaweeks + blength`
```{r}
fit1 = lm(bwt ~ malform + mrace + frace + smoken + gaweeks + bhead, data = bw_df)

broom::tidy(fit1) |> knitr::kable()
broom::glance(fit1) |> knitr::kable()

fit2 = lm(bwt ~ malform + mrace + frace + smoken + gaweeks + bhead + blength, data = bw_df)

broom::tidy(fit2) |> knitr::kable()
broom::glance(fit2) |> knitr::kable()

fit = lm(bwt ~ mrace + frace + smoken + gaweeks + blength + bhead, data = bw_df)

broom::tidy(fit) |> knitr::kable()
broom::glance(fit) |> knitr::kable()
```

## show a plot of model residuals against fitted values based on my model `fit`

```{r}
bw_df |> 
  select(bwt, mrace, frace, smoken, gaweeks, blength, bhead) |> 
  add_predictions(fit) |> 
  add_residuals(fit) |> 
  ggplot(aes(y = resid, x = pred)) + geom_point(alpha = 0.5) +
  labs(x = "fitted value", y = "residuals", title = "residuals against fitted values")
```

* based on this plot, we can see some fitted values have large residuals, so maybe there are outliers when we use model fit for prediction. Since some residuals are pretty large, we cannot say that the model fit is the best fitted model for this dataset. 



